Examining spatial equity and accessibility to a public bicycle share program using a balanced floating catchment area approach

### 2021-03-18 Bike-share-spatial-equity/Bike-share-spatial-equity.Rmd
Revised draft

```{r load-packages, include=FALSE}
library(data.table)
library(dplyr)
library(gdistance)
library(kableExtra)
library(raster)
library(rgdal)
library(sf)
library(sobiEquity)
library(tidyverse)
library(units)
#library(disk.frame)
#library(pycno)
library(readr)
#library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(tinytex)
library(ggplot2)
theme_set(theme_bw())
```

@@

This paper uses the city of Hamilton, located in Ontario, Canada, as a case study. The city launched a public bicycle share program, Hamilton Bike Share, in March 2015 with 115 stations and 750 bicycles [@hamiltonHamiltonBikeShare2015]. Before June 2020, the program was known as SoBi Hamilton. Stations are spaced between 300 and 600 metres apart [@scottRouteChoiceBike2021]. The core service area spans 35 km^2 of the city and _approximately 130,000 can reach a bike share hub within 30 minutes of walking_ [see Figure 1]. This represents roughly one fifth of the population in Hamilton, according to the 2016 Canadian Census. The program was enthusiastically welcomed in the city - within three weeks of launching, 10,000 trips had been made [@hamiltonHamiltonBikeShare2015]. In 2017, Hamilton Bike Share Inc., the non-profit organization that operates the program, initiated an equity program, Everyone Rides Initiative (ERI), to remove barriers that may prevent individuals from accessing bike share in Hamilton. An additional 75 bicycles and 15 stations were added to the program, which expanded it to more disadvantaged areas in the core service area [see Figure 2]. The program also offers subsidized memberships to individuals who identify as low income. A comparable program can be found in Philadelphia [see @caspiBikesharingPhiladelphiaLowerincome2019]. As of June 2020, the bike share program has 900 bikes and 130 stations, and over 26,000 active memberships [@hamiltonHamiltonBikeShare2015].

@@

## Accessibility Calculations at Different Thresholds

@@

### Varying thresholds

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 3 minutes:
```{r}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

All stations:
```{r}
results_3min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

Sanity check:
```{r}
sum(results_3min_conventional$los$los)
sum(results_3min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_3min_all$los$los)
sum(results_3min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 5 minutes:
```{r}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

All stations:
```{r}
results_5min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

Sanity check:
```{r}
sum(results_5min_conventional$los$los)
sum(results_5min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_5min_all$los$los)
sum(results_5min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 10 minutes:
```{r}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

All stations:
```{r}
results_10min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

Sanity check:
```{r}
sum(results_10min_conventional$los$los)
sum(results_10min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_10min_all$los$los)
sum(results_10min_all$accessibility$accessibility)
```

For the system without the equity stations we remove the equity stations from the travel time matrix before calculating accessibility in at most 15 minutes:
```{r}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

All stations:
```{r}
results_15min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

Sanity check:
```{r}
sum(results_15min_conventional$los$los)
sum(results_15min_conventional$accessibility$accessibility)
```

Sanity check:
```{r}
sum(results_15min_all$los$los)
sum(results_15min_all$accessibility$accessibility)
```

Extract results:
```{r, include=FALSE}
los <- rbind(data.frame(results_3min_all$los, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$los, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$los, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$los, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$los, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$los, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$los, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$los, system = "Conventional", threshold = "15 min"))

@@

Add geometry:
```{r}
los <- los %>%
  left_join(sobi_hubs %>%
              dplyr::select(OBJECTID),
            by = c("hub" = "OBJECTID")) %>%
  st_as_sf()

@@

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 3 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "3 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 5 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "5 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 10 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "10 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r, echo=FALSE}
#accessibility plot for current system with equity stations at 15 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "15 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

@@

>>>>

```{r, include=FALSE}
rm(list = ls())
```

```{r load-packages, include=FALSE}
library(data.table)
library(dplyr)
library(gdistance)
library(kableExtra)
library(raster)
library(rgdal)
library(sf)
library(sobiEquity)
library(tidyverse)
library(units)
#library(disk.frame)
#library(pycno)
library(readr)
#library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html
library(tinytex)
library(ggplot2)
library(biscale)
library(cowplot)
```

@@

This paper uses the city of Hamilton, located in Ontario, Canada, as a case study. The city launched a public bicycle share program, Hamilton Bike Share, in March 2015 with 115 stations and 750 bicycles [@hamiltonHamiltonBikeShare2015]. Before June 2020, the program was known as SoBi Hamilton. Stations are spaced between 300 and 600 metres apart [@scottRouteChoiceBike2021]. The core service area spans 35 km^2 of the city and roughly 138,000 can reach a bike share hub within 30 minutes of walking [see Figure 1]. This represents roughly one fifth of the population in Hamilton, according to the 2016 Canadian Census. The program was enthusiastically welcomed in the city - within three weeks of launching, 10,000 trips had been made [@hamiltonHamiltonBikeShare2015]. In 2017, Hamilton Bike Share Inc., the non-profit organization that operates the program, initiated an equity program, Everyone Rides Initiative (ERI), to remove barriers that may prevent individuals from accessing bike share in Hamilton. An additional 75 bicycles and 15 stations were added to the program, which expanded it to more disadvantaged areas in the core service area [see Figure 2]. The program also offers subsidized memberships to individuals who identify as low income. A comparable program can be found in Philadelphia [see @caspiBikesharingPhiladelphiaLowerincome2019]. As of June 2020, the bike share program has 900 bikes and 130 stations, and over 26,000 active memberships [@hamiltonHamiltonBikeShare2015].

@@

## Accessibility at Different Thresholds

@@

### Three Minutes

For the system without the equity stations, we remove the equity stations from the travel time matrix before calculating accessibility in at most 3 minutes walk.

```{r min-conventional, include=FALSE}
results_3min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

```{r min-all, include=FALSE}
results_3min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 3)
```

```{r min-sum-conventional, echo=FALSE}
sum(results_3min_conventional$los$los)
sum(results_3min_conventional$accessibility$accessibility)
```

```{r min-sum-all, echo=FALSE}
sum(results_3min_all$los$los)
sum(results_3min_all$accessibility$accessibility)
```

### Five Minutes

For the system without the equity stations, we remove the equity stations from the travel time matrix before calculating accessibility in at most 5 minutes walk.

```{r med-conventional, include=FALSE}
results_5min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

```{r med-all, include=FALSE}
results_5min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 5)
```

```{r, med-sum-conventional, echo=FALSE}
sum(results_5min_conventional$los$los)
sum(results_5min_conventional$accessibility$accessibility)
```

```{r med-sum-all, echo=FALSE}
sum(results_5min_all$los$los)
sum(results_5min_all$accessibility$accessibility)
```

### Ten Minutes

For the system without the equity stations, we remove the equity stations from the travel time matrix before calculating accessibility in at most 10 minutes.

```{r avg-conventional, include=FALSE}
results_10min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

```{r avg-all, include=FALSE}
results_10min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 10)
```

```{r avg-sum-conventional, echo=FALSE}
sum(results_10min_conventional$los$los)
sum(results_10min_conventional$accessibility$accessibility)
```

```{r, avg-sum-all, echo=FALSE}
sum(results_10min_all$los$los)
sum(results_10min_all$accessibility$accessibility)
```

### Fifteen Minutes

For the system without the equity stations, we remove the equity stations from the travel time matrix before calculating accessibility in at most 15 minutes walk.

```{r max-conventional, include=FALSE}
results_15min_conventional <- ttm_walk %>%
  dplyr::filter(hub_type == "Conventional" &
                  hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

```{r max-all, include=FALSE}
results_15min_all <- ttm_walk %>%
  dplyr::filter(hub_status == "Active") %>%
  b2sfca(threshold = 15)
```

```{r max-sum-conventional, echo=FALSE}
sum(results_15min_conventional$los$los)
sum(results_15min_conventional$accessibility$accessibility)
```

```{r max-sum-all, echo=FALSE}
sum(results_15min_all$los$los)
sum(results_15min_all$accessibility$accessibility)
```

```{r, include=FALSE}
los <- rbind(data.frame(results_3min_all$los, system = "All stations", threshold = "3 min"),
             data.frame(results_3min_conventional$los, system = "Conventional", threshold = "3 min"),
             data.frame(results_5min_all$los, system = "All stations", threshold = "5 min"),
             data.frame(results_5min_conventional$los, system = "Conventional", threshold = "5 min"),
             data.frame(results_10min_all$los, system = "All stations", threshold = "10 min"),
             data.frame(results_10min_conventional$los, system = "Conventional", threshold = "10 min"),
             data.frame(results_15min_all$los, system = "All stations", threshold = "15 min"),
             data.frame(results_15min_conventional$los, system = "Conventional", threshold = "15 min"))

@@

```{r, include=FALSE}
los <- los %>%
  left_join(sobi_hubs %>%
              dplyr::select(OBJECTID),
            by = c("hub" = "OBJECTID")) %>%
  st_as_sf()

@@

```{r figure-6, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) compared between current system with equity stations and the original system without equity stations."}
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "3 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r figure-7, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 5 minutes walk (average threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 5 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "5 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r figure-8, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 10 minutes walk (maximum threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 10 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "10 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

```{r figure-9, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 15 minutes walk (extreme threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 15 minutes walk
ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  geom_sf(data = accessibility %>%
            dplyr::filter(threshold == "15 min"),
       aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1) +
  facet_grid(system ~ threshold)
```

@@

## Accessibility and Median Household Income

```{r, include=FALSE}
# drop geometry
test <- accessibility %>% st_drop_geometry()

# join accessibility to population cells
test <- test %>% left_join(population_50x50, by = "UID")
```

```{r, include=FALSE}
test0 <- population_50x50 %>%
  st_join(hamilton_da_2016 %>% dplyr::select(DA))
```

```{r, include=FALSE}
test <- test0 %>% left_join(test, by = "UID")
```

```{r, include=FALSE}
# object is a combination of accessibility, population_50x50, with DAs
test <- test %>% drop_na()
```

```{r, include=FALSE}
test <- subset(test, select = -c(1)) %>% dplyr::select(-geometry.y)
```

```{r, include=FALSE}
accessibility <- test
```

```{r, include=FALSE}
min_all <- accessibility %>%
            dplyr::filter(threshold == "3 min") %>% 
            dplyr::filter(system == "All stations")

min_con <- accessibility %>%
            dplyr::filter(threshold == "3 min") %>%
            dplyr::filter(system == "Conventional")

avg_all <- accessibility %>%
            dplyr::filter(threshold == "5 min") %>% 
            dplyr::filter(system == "All stations")

avg_con <- accessibility %>%
            dplyr::filter(threshold == "5 min") %>%
            dplyr::filter(system == "Conventional")

max_all <- accessibility %>%
            dplyr::filter(threshold == "10 min") %>% 
            dplyr::filter(system == "All stations")

max_con <- accessibility %>%
            dplyr::filter(threshold == "10 min") %>%
            dplyr::filter(system == "Conventional")

ext_all <- accessibility %>%
            dplyr::filter(threshold == "15 min") %>% 
            dplyr::filter(system == "All stations")

ext_con <- accessibility %>%
            dplyr::filter(threshold == "15 min") %>%
            dplyr::filter(system == "Conventional")

```

```{r, include=FALSE}
# aggregate population and accessibility by DA
min_all <- min_all %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

min_con <- min_con %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

avg_all <- avg_all %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

avg_con <- avg_con %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

max_all <- max_all %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

max_con <- max_con %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

ext_all <- ext_all %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))

ext_con <- ext_con %>%
           group_by(DA) %>%
             summarize(population = sum(population.y),
             accessibility = sum(accessibility)) %>%
           st_join(hamilton_da_2016 %>% dplyr::select(median_total_income))
```

```{r}
min_all <- min_all %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

min_con <- min_con %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

avg_all <- avg_all %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

avg_con <- avg_con %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

max_all <- max_all %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

max_con <- max_con %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

ext_all <- ext_all %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()

ext_con <- ext_con %>%
  st_drop_geometry() %>%
  left_join(hamilton_da_2016 %>% 
              dplyr::select(DA, geometry),
            by = "DA") %>%
  st_as_sf()
```

```{r}
data_min_all <- bi_class(min_all,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_min_con <- bi_class(min_con,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_avg_all <- bi_class(avg_all,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_avg_con <- bi_class(avg_con,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_max_all <- bi_class(max_all,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_max_con <- bi_class(max_con,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_ext_all <- bi_class(ext_all,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()

data_ext_con <- bi_class(ext_con,
                 x = accessibility, 
                 y = median_total_income, 
                 style = "quantile", 
                 dim = 3) %>%
  st_as_sf()
```
```{r figure-10, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) varying by median total income in the current system with equity stations."}
map_min_all <- ggplot() +
  geom_sf(data = data_min_all, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_min_all, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-11, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) varying by median total income in the original system with conventional stations."}
map_min_con <- ggplot() +
  geom_sf(data = data_min_con, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_min_con, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-12, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 5 minutes walk (average threshold) varying by median total income in the current system with equity stations."}
map_avg_all <- ggplot() +
  geom_sf(data = data_avg_all, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_avg_all, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-13, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 5 minutes walk (average threshold) varying by median total income in the original system with conventional stations."}
map_avg_con <- ggplot() +
  geom_sf(data = data_avg_con, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_avg_con, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-14, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 10 minutes walk (maximum threshold) varying by median total income in the current system with equity stations."}
map_max_all <- ggplot() +
  geom_sf(data = data_max_all, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_max_all, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-15, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 10 minutes walk (maximum threshold) varying by median total income in the original system with conventional stations."}
map_max_con <- ggplot() +
  geom_sf(data = data_max_con, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_max_con, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-16, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 15 minutes walk (extreme threshold) varying by median total income in the current system with equity stations."}
map_ext_all <- ggplot() +
  geom_sf(data = data_ext_all, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_ext_all, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r figure-17, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Accessibility at 15 minutes walk (extreme threshold) varying by median total income in the original system with conventional stations."}
map_ext_con <- ggplot() +
  geom_sf(data = data_ext_con, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  bi_theme()

legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "Higher accessibility ",
                    ylab = "Higher income",
                    size = 6)

cowplot::ggdraw() +
  cowplot::draw_plot(map_ext_con, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.05, .6, 0.2, 0.2)
```

```{r, include=FALSE}
sum(population_50x50$population)
sum(min_all$population)
sum(avg_all$population)
sum(max_all$population)
sum(ext_all$population)
sum(min_con$population)
sum(avg_con$population)
sum(max_con$population)
sum(ext_con$population)
```

```{r}
summary(data_max_all)
```

```{r}
demographics_minimum <- rbind(data.frame(min_all %>%  
                                        dplyr::filter(median_total_income >= "38486") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "All Stations"),
             data.frame(min_all %>% dplyr::filter(median_total_income >= "46336") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "All Stations"),
             data.frame(min_all %>% dplyr::filter(median_total_income >= "56640") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "All Stations"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "38288") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "Conventional"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "45696") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "Conventional"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "57664") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "Conventional"))
```

```{r}
demographics_maximum <- rbind(data.frame(avg_all %>%  
                                        dplyr::filter(median_total_income >= "38752") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "All Stations"),
             data.frame(avg_all %>% dplyr::filter(median_total_income >= "47040") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "All Stations"),
             data.frame(avg_all %>% dplyr::filter(median_total_income >= "58912") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "All Stations"),
             data.frame(avg_con %>% dplyr::filter(median_total_income >= "38422") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "Conventional"),
             data.frame(avg_con %>% dplyr::filter(median_total_income >= "45696") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "Conventional"),
             data.frame(avg_con %>% dplyr::filter(median_total_income >= "59264") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "Conventional"))
```

```{r}
demographics_minimum <- rbind(data.frame(min_all %>%  
                                        dplyr::filter(median_total_income >= "38486") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "All Stations"),
             data.frame(min_all %>% dplyr::filter(median_total_income >= "46336") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "All Stations"),
             data.frame(min_all %>% dplyr::filter(median_total_income >= "56640") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "All Stations"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "38288") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 1", system = "Conventional"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "45696") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 2", system = "Conventional"),
             data.frame(min_con %>% dplyr::filter(median_total_income >= "57664") %>% summarise(population = sum(population)), accessibility = "min", income = "Quantile 3", system = "Conventional"))
```

```{r, include=FALSE}
sum(population_50x50$population)
sum(min_all$population)
sum(avg_all$population)
sum(max_all$population)
sum(ext_all$population)
sum(min_con$population)
sum(avg_con$population)
sum(max_con$population)
sum(ext_con$population)
```

```{r table-2, echo=FALSE, fig.align = 'center', out.width = "65%", fig.cap = "Population that can reach at least one bike share station at different thresholds in the current system with equity stations and original system with only conventional stations."}
data_tbl <- data.frame(
  System = c("All Stations", "Conventional"),
  Minimum = c("48,725", "43,238"),
  Average = c("84,655", "75,026"),
  Maximum = c("111,680", "104,298"),
  Extreme = c("118,581", "115,344")
)

kbl(data_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  kable_styling(bootstrap_options = "striped") %>%
  add_header_above(c(" " = 1, "3 Minutes" = 1, "5 Minutes" = 1, "10 Minutes" = 1, "15 Minutes" = 1))
```
